### 2024-01-26 ms.qmd
Updates

## Results

@@

>>>>

## Results
```{r, sampleSize}
#| label: sampleSize
# List with the sample sizes from the database (here("output/databases_clean/data_asso.csv") as the sample size per species and group is the same on each task. We used function sample (see func.R) to estimate the sample size per treatment and species.
source(here("R", "func.R"))
#
specie <- c("delicata", "guichenoti")
groups <- c("Red", "Blue")
hormone <- c("CORT", "Control")
temperature <- c("Cold", "Hot")
#
n_list <- list()
#
for(i in 1:length(specie)){
  for(j in 1:length(groups)){
    for(k in 1:length(hormone)){
      for(l in 1:length(temperature)){
        n_list[[paste(specie[i], groups[j], hormone[k], temperature[l], sep = "_")]] <- sample(specie[i], groups[j], hormone[k], temperature[l])
      }
    }
  }
}
```

Originally, we started with 96 lizards, 48 per species and 12 per treatment per species. However, due to natural mortality (n = 11), no completion of the training stage (n = 1), or no motivation during the learning tasks (n = 3; see above), we ended up with a total of 81 lizards. Final sample sizes per treatment and species are disclosed on @tbl-data.

```{r, models}
#| label: models
# Fitting the model and extraction of posteriors for both types of task and species using fit_m function (see func.r in R folder). The result everytime the function is used is a df with the posteriors of the model. The functions saves the model automatically in output/models; and when the parameter refit = FALSE then the posteriors are extracted from the model previously written instead of fitting the model again each time.
source(here("R", "func.R"))
# A) Associative task. Model formula: FC_associative ~ Associative_Trial*cort*temp + (1 + Associative_Trial|lizard_id)
## A.1) L. delicata
deli_asso_red <- fit_m("asso", "deli", "red", refit = FALSE)
write.csv(deli_asso_red, file= "./output/Checking/deli_asso_red.csv")
deli_asso_blue <- fit_m("asso", "deli", "blue", refit = FALSE)
write.csv(deli_asso_blue, file= "./output/Checking/deli_asso_blue.csv")
## A.2) L. guichenoti
guich_asso_red <- fit_m("asso", "guich", "red", refit = FALSE)
write.csv(guich_asso_red, file= "./output/Checking/guich_asso_red.csv")
guich_asso_blue <- fit_m("asso", "guich", "blue", refit = FALSE)
write.csv(guich_asso_blue, file= "./output/Checking/guich_asso_blue.csv")
# B) Reversal task. Model formula: FC_reversal ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)
## B.1) L. delicata
deli_rev_red <- fit_m("rev", "deli", "red", refit = FALSE)
write.csv(deli_rev_red, file= "./output/Checking/deli_rev_red.csv")
deli_rev_blue <- fit_m("rev", "deli", "blue", refit = FALSE)
write.csv(deli_rev_blue, file= "./output/Checking/deli_rev_blue.csv")
## B.2) L. guichenoti
guich_rev_red <- fit_m("rev", "guich", "red", refit = FALSE)
write.csv(guich_rev_red, file= "./output/Checking/guich_rev_red.csv")
guich_rev_blue <- fit_m("rev", "guich", "blue", refit = FALSE)
write.csv(guich_rev_blue, file= "./output/Checking/guich_rev_blue.csv")
```

```{r, resultsRev}
# Rename some of the posteriors and make new estimates for the learning rate for the Reversal task doing the same thing we did in the chunk above.
## 1) L. delicata
### Group = red
drr_CORTCold <- deli_rev_red$b_trial_reversal
drr_ControlCold <- (deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$b_trial_reversal)
drr_CORTHot <- (deli_rev_red$'b_trial_reversal:tempHot' + deli_rev_red$b_trial_reversal)
drr_ControlHot <- (deli_rev_red$'b_trial_reversal:cortControl:tempHot' + deli_rev_red$b_trial_reversal + deli_rev_red$'b_trial_reversal:cortControl' + deli_rev_red$'b_trial_reversal:tempHot')
### Group = blue
drb_CORTCold <- deli_rev_blue$b_trial_reversal
drb_ControlCold <- (deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$b_trial_reversal)
drb_CORTHot <- (deli_rev_blue$'b_trial_reversal:tempHot' + deli_rev_blue$b_trial_reversal)
drb_ControlHot <- (deli_rev_blue$'b_trial_reversal:cortControl:tempHot' + deli_rev_blue$b_trial_reversal + deli_rev_blue$'b_trial_reversal:cortControl' + deli_rev_blue$'b_trial_reversal:tempHot')
## 2) L. guichenoti
### Group = red
grr_CORTCold <- guich_rev_red$b_trial_reversal
grr_ControlCold <- (guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$b_trial_reversal)
grr_CORTHot <- (guich_rev_red$'b_trial_reversal:tempHot' + guich_rev_red$b_trial_reversal)
grr_ControlHot <- (guich_rev_red$'b_trial_reversal:cortControl:tempHot' + guich_rev_red$b_trial_reversal + guich_rev_red$'b_trial_reversal:cortControl' + guich_rev_red$'b_trial_reversal:tempHot')
### Group = blue
grb_CORTCold <- guich_rev_blue$b_trial_reversal
grb_ControlCold <- (guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$b_trial_reversal)
grb_CORTHot <- (guich_rev_blue$'b_trial_reversal:tempHot' + guich_rev_blue$b_trial_reversal)
grb_ControlHot <- (guich_rev_blue$'b_trial_reversal:cortControl:tempHot' + guich_rev_blue$b_trial_reversal + guich_rev_blue$'b_trial_reversal:cortControl' + guich_rev_blue$'b_trial_reversal:tempHot')
```

```{r, tbl-data}
#| label: tbl-data
#| tbl-cap: "Estimates of Associative learning slope for all the different treatments per each task, species and group. Mean shows the aritmetic mean of the estimates obtained from the posteriors of the model, and 95% CI indicates the 95% confidence interval of the mean. All p-values were obtained using pmcmc and test the hypothesis that the mean is equal to zero. In bold, those values that are significant (p-value <0.05)"
#
############################## CREATING BIG DF FOR TABLE ##############################
# Building the vectors for titles of rows and columns
specie <- c("L. delicata", "L. guichenoti")
groups <- c("Red", "Blue")
test <- c("Associative", "Reversal")
treatments <- c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")
values <- c("Mean", "95% CI", "p-value")
# Building the vectors for estimated means, co.intervals(95%), and p-values for the slopes obtained from posteriors. p-values are obtained using pmcmc function (see func.R), assuming a two-tailed test that testes the hypothesis that the value (slopes in this case) is 0.
#First get estimates for both tasks
estimates_asso <- list(
  dar_CORTCold, dar_ControlCold, dar_CORTHot, dar_ControlHot, 
  dab_CORTCold, dab_ControlCold, dab_CORTHot, dab_ControlHot, 
  gar_CORTCold, gar_ControlCold, gar_CORTHot, gar_ControlHot, 
  gab_CORTCold, gab_ControlCold, gab_CORTHot, gab_ControlHot
)
#
estimates_rev <- list(
  drr_CORTCold, drr_ControlCold, drr_CORTHot, drr_ControlHot, 
  drb_CORTCold, drb_ControlCold, drb_CORTHot, drb_ControlHot, 
  grr_CORTCold, grr_ControlCold, grr_CORTHot, grr_ControlHot, 
  grb_CORTCold, grb_ControlCold, grb_CORTHot, grb_ControlHot
)
# Then get the mean, co.intervals(95%), and p-values
asso_mean <- format_dec(sapply(estimates_asso, mean), 3)
asso_interval_025 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.025)), 3)
asso_interval_975 <- format_dec(sapply(estimates_asso, function(x) quantile(x,0.975)), 3)
asso_intervals <- paste(asso_interval_025, asso_interval_975, sep = " , ")
asso_pvalue <- format_dec(sapply(estimates_asso, pmcmc), 3)
#
rev_mean <- format_dec(sapply(estimates_rev, mean), 3)
rev_interval_025 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.025)), 3)
rev_interval_975 <- format_dec(sapply(estimates_rev, function(x) quantile(x,0.975)), 3)
rev_intervals <- paste(rev_interval_025, rev_interval_975, sep = " , ")
rev_pvalue <- format_dec(sapply(estimates_rev, pmcmc), 3)
#
# Building the df for the associative
asso_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(asso_mean, each = 1),
  CI = rep(asso_intervals, each = 1),
  PValue = rep(asso_pvalue, each = 1),
  Task = rep("Associative", length(asso_mean))
)
# Building the df for the reversal
rev_df <- data.frame(
  Specie = rep(specie, each = length(groups) * length(treatments)),
  Group = rep(rep(groups, each = length(treatments)), times = length(specie)),
  Treatment = rep(rep(treatments, each = 1), times = length(groups) * length(specie)),
  Mean = rep(rev_mean, each = 1),
  CI = rep(rev_intervals, each = 1),
  PValue = rep(rev_pvalue, each = 1),
  Task = rep("Reversal", length(rev_mean))
)
# Joining both dfs
table_data <- rbind(asso_df, rev_df)
table_data[, sapply(table_data, is.numeric)] <- lapply(table_data[, sapply(table_data, is.numeric)], function(x) format(x, scientific = FALSE))
#
############################## ADDING SAMPLE SIZE TO DF FOR TABLE ##############################
# Make n_list into a df
n_df <- as.data.frame(do.call(rbind, n_list)) %>%
  rename("n" = V1) %>%
  rownames_to_column("model") %>%
  separate(model, into = c("Specie", "Group", "cort", "temp"), sep = "_") %>%
  unite("Treatment", c("cort", "temp"), sep = "-") %>%
                  labels = c(delicata = "L. delicata", guichenoti = "L. guichenoti")),
        Treatment = factor(Treatment,
                   levels = c("CORT-Cold", "Control-Cold", "CORT-Hot","Control-Hot")))
# Merge both dfs, put sample size together with the treatment, and organize the new df to make it look like the table
new_table_data <- merge(table_data, n_df) %>%
  rename('p-value' = 'PValue', '95% CI' = 'CI') %>% #Change the names of the columns for the table
  pivot_wider(names_from = Task, values_from = c(Mean, `95% CI`, `p-value`)) %>% # to split between Asociative and Reversal
  select(Specie, Group, Treatment, Mean_Associative, `95% CI_Associative`, `p-value_Associative`, Mean_Reversal, `95% CI_Reversal`, `p-value_Reversal`, n) %>% #To order the columns in the way I want for the table
                  levels = c("L. delicata", "L. guichenoti")),
        Group = factor(Group,
                  levels = c("Red", "Blue")),
        Treatment = factor(Treatment, 
                  levels = c("CORT-Cold", "Control-Cold", "CORT-Hot", "Control-Hot")))%>%
  arrange(Specie, Group, Treatment) %>% # To arrange the rows the way I want
  unite("Treatment", c("Treatment", "n"), sep = " (n = ") %>%
  mutate(Treatment = paste0(Treatment, ")"))
write.csv(new_table_data, file= "./output/Checking/new_table_data.csv")
#
############################## MAKING THE TABLE ##############################
## Table format
set_flextable_defaults(
 font.family = "Times New Roman",
 fint.size = 10)
# Split the table_data df by task
real_table <- flextable(new_table_data) %>%
    bold(~ `p-value_Associative` < 0.05, ~ `p-value_Associative` + Mean_Associative + `95% CI_Associative`) %>%
    bold(~ `p-value_Reversal` < 0.05, ~ `p-value_Reversal` + Mean_Reversal + `95% CI_Reversal`) %>%
    add_header_row(values = c("", "Associative task", "Reversal task"), colwidths = c(3, 3, 3)) %>%
    set_header_labels(Mean_Associative = "Mean",
                      `95% CI_Associative` = "95% CI",
                      `p-value_Associative` = "p-value",
                      Mean_Reversal = "Mean",
                      `95% CI_Reversal` = "95% CI",
                      `p-value_Reversal` = "p-value") %>%
    italic(j = 1, italic = TRUE, part = "body") %>% # To have names od species in italics
    flextable::compose(i = c(2:8,10:16), j = 1, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the first column
    flextable::compose(i = c(2:4,6:8,10:12,14:16), j = 2, value = as_paragraph(""), part = "body") %>% # To remove some of the values in the second column
    hline(i = c(4,12), j = c(2:9), part = "body") %>% # To make some horizontal lines
    hline(i = c(8), j = c(1:9), part = "body") %>% # To make some horizontal lines
    vline(i = (1:16), j = c(3,6), part = "body") %>% # To make some vertical lines on body
    vline(j=c(3,6), part = "header") %>% # To make some vertical lines on header
real_table

In the reversal tesk, we decided to pool both groups 'Blue' and 'Red' to make the between-treatment comparisons for two reasons. First, we did not find any significant differences caused group on the learning slopes (_L. delicata_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`, p = `r format_dec(pmcmc((drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drr_CORTCold + drr_ControlHot)), 3)`; _L. guichenoti_: 'Blue' learning slope - 'Red'learning slope = `r format_dec(mean((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`, p = `r format_dec(pmcmc((grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTCold + grr_ControlHot)), 3)`). Second, we were expecting an acquired bias towards the colour assigned in the associative task as part of the design. Nonetheless, we did not find any significant differences between treatments in _L. delicata_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlCold + drb_ControlCold) - (drr_CORTCold + drb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_CORTHot + drb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlHot + drb_ControlHot) - (drr_ControlCold + drb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drb_CORTHot) - (drr_CORTCold + drb_CORTCold)), 3)`) (see @Figdeli) or _L. guichenoti_ ('Control-Cold' - 'CORT-Cold' = `r format_dec(mean((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlCold + grb_ControlCold) - (grr_CORTCold + grb_CORTCold)), 3)`; 'Control-Hot' - 'CORT-Hot' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`, p- value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_CORTHot + grb_CORTHot)), 3)`; 'Control-Hot' - 'Control-Cold' = `r format_dec(mean((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlHot + grb_ControlHot) - (grr_ControlCold + grb_ControlCold)), 3)`; 'CORT-Hot' - 'CORT-Cold' = `r format_dec(mean((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grb_CORTHot) - (grr_CORTCold + grb_CORTCold)), 3)`) (see @Figguich). We did not find any effect of incubation temperature (_L. delicata_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((drr_CORTHot + drr_ControlHot + drb_CORTHot + drb_ControlHot) - (drr_CORTCold + drr_ControlCold + drb_CORTCold + drb_ControlCold)),3)`; _L. guichenoti_: 'Hot' learning slope - 'Cold' learning slope = `r format_dec(mean((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`, p-value = `r format_dec(pmcmc((grr_CORTHot + grr_ControlHot + grb_CORTHot + grb_ControlHot) -(grr_CORTCold + grr_ControlCold + grb_CORTCold + grb_ControlCold)),3)`) or the hormone (_L. delicata_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((drr_ControlCold + drr_ControlHot + drb_ControlCold + drb_ControlHot) - (drr_CORTCold + drr_CORTHot + drb_CORTCold + drb_CORTHot)), 3)`;_L. guichenoti_: 'Control' learning slope - 'CORT' learning slope = `r format_dec(mean((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`, p-value = `r format_dec(pmcmc((grr_ControlCold + grr_ControlHot + grb_ControlCold + grb_ControlHot)-(grr_CORTCold + grr_CORTHot + grb_CORTCold + grb_CORTHot)), 3)`) when groups were pooled. Finally, we did not find any significant differences when we compared the estimated slopes between species for the reversal task (_L. delicata_ - _L. guichenoti_ = `r format_dec(mean((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`, p-value = `r format_dec(pmcmc((drr_CORTCold + drr_ControlCold + drr_CORTHot + drr_ControlHot + drb_CORTCold + drb_ControlCold + drb_CORTHot + drb_ControlHot) - (grr_CORTCold + grr_ControlCold + grr_CORTHot + grr_ControlHot + grb_CORTCold + grb_ControlCold + grb_CORTHot + grb_ControlHot)), 3)`).

@@

#### Checking the models plots  
```{r, residuals}
# Chunk for calling all the models and making the residuals
# Associative task
## L. delicata
### Red
mod_dar <- readRDS(here("output/models/asso_deli_red.rds"))
resid_dar <- residuals(mod_dar)
### Blue
mod_dab <- readRDS(here("output/models/asso_deli_blue.rds"))
resid_dab <- residuals(mod_dab)
## L. guichenoti
### Red
mod_gar <- readRDS(here("output/models/asso_guich_red.rds"))
resid_gar <- residuals(mod_gar)
### Blue
mod_gab <- readRDS(here("output/models/asso_guich_blue.rds"))
resid_gab <- residuals(mod_gab)
# Reversal task
## L. delicata
### Red
mod_drr <- readRDS(here("output/models/rev_deli_red.rds"))
resid_drr <- residuals(mod_drr)
### Blue
mod_drb <- readRDS(here("output/models/rev_deli_blue.rds"))
resid_drb <- residuals(mod_drb)
## L. guichenoti
### Red
mod_grr <- readRDS(here("output/models/rev_guich_red.rds"))
resid_grr <- residuals(mod_grr)
### Blue
mod_grb <- readRDS(here("output/models/rev_guich_blue.rds"))
resid_grb <- residuals(mod_grb)
```

Model formula for the reversal task is:  
Choice  ~ trial_reversal*cort*temp + (1 + trial_reversal|lizard_id)  
Plots for the different models of the associative task:  
1.- _L. delicata_  
  1.a.- Red  
```{r , plotmod_drr, out.width="70%", fig.align="center"}
plot(mod_drr)
```

  1.b.- Blue  
```{r , plotmod_drb, out.width="70%", fig.align="center"}  
plot(mod_drb)
```

2.- _L. guichenoti_  
  2.a.- Red  
```{r , plotmod_grr, out.width="70%", fig.align="center"}
plot(mod_grr)
```

  2.b.- Blue  
```{r , plotmod_grb, out.width="70%", fig.align="center"}
plot(mod_grb)
```
